Species motif participation provides unique information about species risk of extinction

### 2022-06-08 manuscript/response_PRSB_round2.tex

        \us
        We are not convinced that collecting data on trophic level is easier than collecting entire food web data. If the Reviewer is referring to stable-isotope trophic levels, then we must point out that calculating trophic levels requires a good background understanding of the food sources for the focal species and rates of incorporation of different isotopes across trophic levels (and how these may differ between taxa). For a large community (e.g., 100 species), it could well be simpler to at least approximate the entire food web (giving trophic level, degree, and motifs simultaneously) than to calculate stable isotope-based trophic levels for each species. Indeed, the larger set of empirical food webs than comprehensive stable-isotope data suggests that food webs are easier to estimate at scale than trophic levels alone.

        To address this concern, we have added a brief statement in the introduction to make it clear to readers that researchers who have enough information to calculate in-degree and trophic level can also calculate motifs. Motif roles, therefore, provide additional information than degree and trophic level alone without requiring extra sampling.

        \them
        Third, I realize that this technique is already published on, however, some of the limitations of the technique may impact the utility of the results. I think the study would be greatly enhanced by a smaller comparison of a more traditional dynamic food web model approach and this approach – even just placed in the appendix. I note that the authors rely quite heavily on a paper on BioRxivs which has done something similar, so anticipate this wouldn’t be that difficult, but would be very convincing.

        \us We agree that a comparison between Bayesian and dynamical simulations could be interesting, however performing such a comparison is beyond the scope of this manuscript. As the Reviewer notes, a subset of the authors of this manuscript are working on an analysis of the relationships between motif participation and persistence in a dynamical framework (currently under review). We do not believe that including the material from this second paper in an appendix is justified when there are already a published comparisons of extinctions simulated by Bayesian and dynamical approaches (\citealp[]{Eklof2013a}). That paper focused on comparing the number and identity of secondary extinctions between the two approaches, and additionally evaluated which functional response of consumers to loss of prey that most accurately predicted the secondary extinctions from a fully-fledged dynamical model. We use this information in the current manuscript.

@@

        \us See above. The only methodological tie between the two manuscripts was when simulating the networks. The underlying software used is peer-reviewed~\citep{Delmas2017} and has been used in several other published studies, as well as the pre-print originally referenced (currently under review). To avoid the mis-perception that we rely on unproven methods, we now present the full network-generation methods in an Appendix.

    \subsubsection*{Minor comments}

        \them
        Line 65 – 70: I agree that in-degree and trophic level may be coarse description of a species’ role within its food-web. However, the examples provided do not convince me that species’ roles based on motifs are a more nuanced way of describing a species’ role within its food web since the two examples provided refer to the influence disturbance may have on a basal resource – an influence that I’m not convinced a species’ role based on motifs will impact.

        \us
        We have revised the introduction throughout, and have taken care to clearly link the extra information given by motifs to our intuition about extinction risk. In particular, motifs give information about a consumer's prey that may reflect the prey's risk of extinction and thereby the predators' risk (e.g., a predator consuming generalist prey will participate in many three-species chain motifs). We have also added a conceptual figure illustrating the four motifs and explaining how different motifs imply different pathways for bottom-up disturbances.

@@

        \them
        Line 87 – 89: similar to comment above – I think the idea of global and local food web properties has not been introduced at all, nor the classic food web properties (which ones are being measured? Only the four specific motifs thought to contribute to stability are mentioned).

        \us We have revised the introduction to more clearly introduce these concepts.

        \them
        Line 92 – 95: this seems to be a flaw to the Bayesian network simulations – we know that mutual predation happens in food webs – especially if the food web isn’t explicitly size structured, and when nested within the broader food web context these feedback loops may have stabilizing, or destabilizing effects.
        \smallskip
        Line 149 – 152: these seem like large limitations to me.

        \us In revising the introduction, we now introduce the Bayesian network approach and its limitations in the methods section. We want to stress that the Reviewer is clearly correct that these are important limitations and we have further clarified that in the text. At the same time we have clarified the clear advantages that the Bayesian network approach provides.
        First, we clearly state that the lack of top-down effects in the Bayesian networks is a limitation. However, we also point out that in prior studies Bayesian simulations have captured most secondary extinctions identified using dynamical simulations. This suggests that, in many cases, bottom-up effects are the main drivers of secondary extinction. We also emphasize the major gain in computational speed when using the Bayesian approach, which allows us to investigate different levels of disturbance rather than simple species removals that needs to be repeated numerous times. Finally, we have also added text to clarify how the process of making the food webs acyclic can affect the results, but that food web robustness rarely is affected.

### 2022-06-10 manuscript/reorder_main_text.tex
Updates from Overleaf

    %\subsection*{2 Global and local network properties are known to affect stability}
    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has for long been vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances. Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
    The undisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}. However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004, Vieira2015}. Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.

    Species properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator). This can be referred to local properties in the network.
    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and secondary extinctions due to reduced plant populations are particularly detrimental \citep{Curtsdotter2011}. %including secondary extinction due to reduced plant populations.
    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
    This since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct.

    %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
    It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}. 
    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree.
    These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}. 
    By allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.

@@

% Severity of disturbance
    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
    A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
    Empirical studies have shown that disturbances to plant species can also affect some consumers more strongly than others  \citep{byrnes2011climate}, and that the strength of disturbance also affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}.

@@

		% BN limitations
		Modelling species extinctions in food webs using Bayesian networks has two main limitations. First, in Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances. However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}. Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems. For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}. Second, due to the strict bottom-up approach of Bayesian networks the networks must be acyclic. Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited. Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.

@@

		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function "nichemodel" within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}. 
		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04). 
		For full details, see \emph{Appendix S1}.
		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.

@@

    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and STL), 3) how network mean persistence (the average persistence of consumers) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
    	See Appendix \emph{3.1 - 3.5} for details. 
    	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.

@@

        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain increased, but persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel). 
        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Table S1, \emph{Appendix S4}).
        The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.

@@

    \subsection*{Network mean persistence and motif profiles}
        % Updated to use means instead of all species
        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}). 
        Average persistence was higher in networks with higher proportions of the chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
        Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).

@@

    % Moving to network context
    As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
    In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
    Connectance did, however, affect which motifs were most common within a network. 
    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
    Global-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.

@@

    % High C webs and omnivory
    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\citep{McLeod2021}.
    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
    In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.
    Although species in our model cannot swap prey, the reduced extinction risk for omnivorous species resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.

@@

    In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
    Our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
    There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.pdf}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This web contains examples of the four motifs present in an acyclic Bayesian network: apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species.
        Note that the nodes participate in more motifs than the ones highlighted. \textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently. Inserted table shows the number of times A and B, respectively, participate in the four different motifs.}
    \label{fig:concept}
    \end{figure}

@@

>>>>

    %\subsection*{2 Global and local network properties are known to affect stability}
    How disturbances, such as population decline or extinction, lead to loss of other species via direct and indirect effects (i.e., secondary extinctions) has for long been vibrant area of research \citep{Santos2021,curtsdotter2011robustness, dunne2009cascading, Eklof2006}.
    Several global (whole-network) as well as local (single-species) properties are known to influence how species respond to disturbances. Two well-known, influential global properties are network size, the number of species present in the network, and network connectance, the number of realized feeding interactions.
    Based on simulations of random networks, large networks were initially thought to be less stable~\citep{May1972}.
    The indisputable presence of large, stable, empirical networks has spurred research into how other structural characteristics, including connectance, may allow large numbers of species to persist~\citep{Dunne2002}.
    Highly-connected networks may be more resistant to secondary extinctions because consumers, on average, have access to a larger number of prey species and can switch from a declining or extinct prey to another resource~\citep{Dunne2002, Eklof2006,Baumgartner2015}. However, a well-connected network also includes more pathways through which a disturbance can spread~\citep{Dunne2004, Vieira2015}. Depending on the strength and arrangements of interactions, this means that species may be impacted through several paths, amplifying the original disturbance and causing more secondary extinctions.
    The net effect of connectance on extinction therefore likely depends on finer-scale network structure, as well as the type and severity of disturbance.

    Species properties known to affect species' response to disturbances include trophic level (a measure of the path length between a consumer and a basal resource) and in-degree (number of prey items of a predator). This can be referred to local properties of the network.
    Species with high trophic levels (long paths to basal resources) and low in-degree (few prey) are generally at higher risks of secondary extinction \citep{binzer2011susceptibility, Eklof2006}, and secondary extinctions due to reduced plant populations are particularly detrimental \citep{Curtsdotter2011}. %including secondary extinction due to reduced plant populations.
    However, in-degree and trophic level are both relatively coarse descriptions of a species' role in its food web~\citep{Cirtwill2018FoodWebs}. 
    This since, as well as the properties of the consumer itself, a consumer's risk of secondary extinction after a disturbance to plants depends on the extinction risk of its prey.
    For example, a generalist consuming many specialist (and therefore high-risk) prey may, for example, be at greater risk of extinction than a specialist which consumes a highly generalist resource that is more unlikely to go extinct.

    %\subsection*{3 Motif participation may similarly affect species' particular extinction risk}
    It is possible to capture information about how a consumer and its prey fit together within a community by defining participation in different motifs -- unique combinations of small numbers of interacting species~\citep{Stouffer2007,Stouffer2012}. 
    Such motif profiles of species are easily calculated once we know the network structure and require no more information than is used to calculate degree or trophic level.
    These `motif participation' roles capture information about species' direct as well as indirect interactions, incorporating larger-scale network structure into species-level descriptions~\citep{Cirtwill2015a,Cirtwill2018FoodWebs}. We refer to these motifs as meso-scale structures, to emphasise their role as structures between global (network) and local (species) properties.  A species' motif participation profile shows the second step indirect interactions the species are affected by, and thereby capture important structures that neither global or local properties do.  
    Therefore, by allowing researchers to distinguish between, for example, consumers of specialists and generalists, motif participation may complement degree and trophic level to more fully predict species' extinction risk.

@@

% Severity of disturbance
    As mentioned above, the severity of a disturbance may interact with network structure to affect secondary extinctions.
    A mild disturbance (e.g., leading to a small increase in the probability of basal resources going extinct) is intuitively likely to cause fewer secondary extinctions than a severe disturbance, and the difference in effects can depend on network structure~\citep{Baumgartner2015}.
    Empirical studies have shown that disturbances to plant species can affect some consumers more strongly than others  \citep{byrnes2011climate}, that the strength of disturbance affects which consumer species are most strongly affected~\citep{detmer2021variation,carnell2020more}, and that the resilience and return time of the ecological community can be strongly influenced by disturbance severity \citep{rydgren2004disturbance}. The severity disturbance therefore clearly has an effect on how the ecological community will respond.

@@

		% BN limitations
		Modelling species extinctions in food webs using Bayesian networks has two main limitations. First, in Bayesian networks consumers do not affect the extinction probabilities of their resources, and as such we can here only capture bottom-up effects of disturbances. However, Bayesian networks has been shown to capture the majority of secondary extinctions produced fully-fledged dynamical models \citep{Eklof2013}. \cite{Eklof2013} analyzed which functional form of consumers response to loss of resources that most accurately predicted the secondary extinctions realized in the dynamical model, and we here use that functional form in our analyzes.  Moreover, bottom-up effects are often the main force of secondary extinctions in empirical systems. For example, bottom-up effects regulate zoo-plankton communities in freshwater lakes, regulating zoo-plankton communities \citep{li2020bottom} and control abundances and biomass of vertebrate species in several terrestrial systems \citep{Dobson2009food, Mduma1999food, Georgiadis2007}. Second, due to the strict bottom-up approach of Bayesian networks the networks must be acyclic. Fortunately, \cite{Allesina2009functional} proved that cyclic and acyclic versions of a network have the same properties in terms of secondary extinctions and the effect on network robustness are very limited. Thus, making the networks acyclic in order to use the Bayesian network approach will not strongly affect extinction scenarios.

@@

		We generated a realistic set of simulated networks based on the niche model~\citep{Williams2000,Stouffer2007} using the function ``nichemodel'' within the Julia~\citep{Bezanson2017julia} package \emph{BioEnergeticFoodWebs}~\citep{bioenergfw}. 
		To capture a range of plausible network architectures, we simulated networks with sizes (S) ranging from 50 to 100 species (in steps of 10) and connectances (C) ranging from 0.02 to 0.18 (in steps of 0.04). 
		For full details and a comparison to empirical networks, see \emph{Appendix S1}.
		We removed any biologically unlikely networks (Appendix \emph{S1}) and replaced them with new simulated networks, repeating the process until we obtained 100 suitable networks in each combination of size and connectance.

@@

    	We tested 1) how species persistence varied with motif participation, 2) how motif participation varied with other network properties (in-degree and trophic level, STL), 3) how network mean persistence (the average persistence of all consumers in the network) varied with network motif profiles, 4) how network motif profiles varied with global structural properties (S and C), and 5) how species persistence and network mean persistence varied with other network properties (S, C, in-degree, and trophic level).
        Where applicable, we also consider how interactions between network structure and disturbance level might affect extinction risk.
    	See Appendix \emph{3.1 - 3.5} for details. 
    	All analyses are briefly described below and were performed using the R~\citep{R} packages \emph{lmerTest}~\citep{lmerTest}, \emph{MuMIn}~\citep{MuMIn}, and \emph{vegan}~\citep{vegan}.

@@

        At high levels of disturbance, a species' persistence decreased as the proportion of the three-species chain in its motif profile increased. Persistence increased as the proportion of any other motif increased (Figure~\ref{fig:prop_lmer_all}; lower lines in each panel). 
        This means that increased disturbance strengthened the effects of the three-species chain motif and changed the direction of the effect of the apparent competition motif  (Figure~\ref{fig:prop_lmer_all}, Table S1, \emph{Appendix S4}).
        The relationship between omnivory and persistence was still positive but weaker at higher levels of disturbance, and the relationship between persistence and direct competition was largely unaffected by disturbance.

@@

    \subsection*{Network mean persistence and motif profiles}
        % Updated to use means instead of all species
        The overall motif profile of a network was related to the average persistence of consumers ($F_{1,2998}$=645, $p$\textless0.001), although this result must be considered tentative due to non-homogeneous distribution of roles (\emph{Appendix S8}). 
        Average persistence was higher in networks with higher proportions of the three-species chain, direct competition, and apparent competition motifs (Table S4, \emph{Appendix S8}). 
        Conversely, average persistence was lower in networks with higher proportions of the omnivory motif. 
        These trends were consistent across levels of disturbance (Fig.~\ref{fig:motif_profile_persistence}, Table S4), \emph{Appendix S8}).% Network persistence varies between networks with different dispersion of motif profiles ($F_{134,2865}$=2.19$\times10^{24}$, $p$=\textless0.001). For example, networks with more variable motif profiles have higher persistence ($\beta$=0.183, $p$\textless0.001).

@@

    % Moving to network context
    As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
    In fact, a network's motif profile had a stronger relationship to mean persistence compared to both network size and connectance. 
    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
    Connectance did, however, affect which motifs were most common within a network. 
    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
    Global-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.

@@

    % High C webs and omnivory
    Supporting this possibility, species participate more frequently in the omnivory motif in high-connectance networks and the relationship between persistence and participation in the omnivory motif is more strongly positive (Fig.~\ref{fig:density_prop_C}) as in~\cite{McLeod2021}.
    These networks contain many pathways by which a bottom-up disturbance can propagate through the network, leading to stronger effects of disturbance on consumers (Table S6; \emph{Appendix S10}).
    In such a network, it may be especially beneficial for a species to feed on multiple trophic levels and, in a sense, spread its risk of losing prey to bottom-up effects.
    Although species in our model cannot swap prey, the reduced extinction risk for omnivorous species resembles empirical studies showing that adaptive omnivory can increase the permanence of three-species chains~\citep{Fagan1997, Kvrivan2005, AbramsFung2010}.

@@

    In summary, our results show that participating in different three-species motifs is related to a species' risk of extinction following a bottom-up disturbance, and that these relationships synthesize information provided by other, simpler measures of species roles.
    Most interestingly, our results also underscore the importance of considering the level of disturbance when identifying which species are most at risk of extinction.
    There were strong interactions between the level of disturbance and the relationships between persistence and multiple measures of species roles (motifs and in-degree).
    These simulated results echo other warnings about `tipping points' where previously beneficial traits can become risky~\citep{Latty2019,Golubski2016,Tylianakis2014}.
    For example, at low levels of disturbance a high degree may be protective since species can compensate if one or a few prey are lost. 
    At high disturbance, however, a high-degree species is likely to be affected by disturbances through many paths and have lower persistence.
    These disturbance effects mean that efforts to define a set of target species for conservation interactions must take the current (and future) level of disturbance the community experiences into account.

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=1.0\textwidth]{manuscript/figures/concept_fig_ver2.png}
        \caption{Conceptual figure of a toy food web. \textbf{a)} This food web contains examples of the four motifs (present in an acyclic directed networks): apparent competition (blue), three-species chain (purple), omnivory (green), and direct competition (yellow). These motifs capture different ways in which disturbances to basal resources can propagate through the network. For example, a disturbance to the bottom species in a three-species chain will have a direct effect on the middle species but only indirectly affect the top species while a disturbance to the bottom species in an omnivory motif will have a direct effect on the middle species and direct and indirect effects on the top species. 
        Note that the nodes participate in more motifs than the ones highlighted. \textbf{b)} In the same food web, species with similar degree and trophic level can have different motif participation; here species A (light grey node with links in the same motif colored light grey) participates in only two chains while species B (dark grey node with its direct consumption links colored dark grey) participates in chains more than tree times more frequently. Inserted table shows the number of times A and B, respectively, participate in the four different motifs. This example highlight that species A may then have a higher risk of going secondary extinct due to its dependence on specialist prey species, compared to species B who depends om generalist prey species. This relationships are captured by the species motif profiles, but not their trophic level or degree.}
    \label{fig:concept}
    \end{figure}

@@

        \begin{figure}[hb!]
        \centering
        \includegraphics[width=0.85\textwidth]{figures/persistence_vs_motifs.eps}
        \caption{The effect of proportion of the role (x-axis) made up by various motifs (columns) on persistence (y-axis). The effect of participating in each motif is based on the fixed effects in Equation~\ref{propreq}. The different colored lines indicate the probability of extinction of basal species, from $\pi_{disturbed} = 0.1$ (top, purple; no disturbance) to $\pi_{disturbed} = 0.5$ (bottom, yellow; high disturbance). 95\% confidence intervals for each line are shown in grey. Note that omnivory made up a smaller proportion of species' roles than other motifs; lines are plotted over the observed ranges of motif participation.}
    \label{fig:prop_lmer_2}
    \end{figure}
